\section{Introduction}
The work presented in this report reflects the assigned task of computing optimizing behavior of a movable arm actuated by two rotors. The optimal path from an initial state to the desired state is computed, both in one and two dimensions. In both cases state feedback is investigated as a way of eliminating drift from the optimal path. Lastly, a nonlinear constraint is added to the optimization problem. Additional constraints are also investigated as a way of limiting nonlinear effects.

The report is organized as follows: Section \ref{text:problem_description} contains a closer description of the lab setup including prelimenary mathematical modeling. Section \ref{text:problem1} describes PID-tuning of the plant as well as model parameter estimation to improve the mathematical models introduced in section \ref{text:problem_description}. In section \ref{text:problem2} open loop optimal control of the travel trajectory is introduced. Section \ref{text:problem3} compares the results of augmenting the basic PID control layer with an LQR based on the optimal input and trajectories calculated in \ref{text:problem2}. In section \ref{text:problem4} an optimal elevation trajectory is also considered, imposed with a non-linear inequality constraint yielding a non-linear optimization problem. Appendix \ref{text:matlab} contains MATLAB scripts used in the work, where scripts used to generate data plots are omitted for brevity. Appendix \ref{text:simulink} contains diagrams of the relevant Simulink models.
